---
title: "TidyTemplate"
date: 2026-01-10
output: html_document
---

# TidyTuesday

Join the Data Science Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(RColorBrewer)
library(cowplot)
library(scales)
library(purrr)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2025-06-03")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
map(tt, glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
gutenberg_authors <- tt$gutenberg_authors
gutenberg_languages <- tt$gutenberg_languages
gutenberg_metadata <- tt$gutenberg_metadata
gutenberg_subjects <- tt$gutenberg_subjects

gutenberg <- gutenberg_languages |> 
        full_join(
                gutenberg_metadata,
                by=join_by(gutenberg_id, language),relationship = "many-to-many"
        ) |> 
        full_join(
                gutenberg_subjects,
                by=join_by(gutenberg_id),relationship = "many-to-many"
        ) |> 
        full_join(
                gutenberg_authors,
                by=join_by(gutenberg_author_id, author),relationship = "many-to-many"
        )

top_langs <- gutenberg_languages %>%
  count(language, sort = TRUE) %>%
  head(10) %>%  # adjust this number
  pull(language)

all_langs <- c("English","French","Finnish",
                                "German","Italian","Dutch",
                                "Spanish","Portuguese","Hungarian",
                                "Chinese")


# Prepare data - filter out single-book combinations
df_upset <- gutenberg_languages %>%
  filter(language %in% top_langs) %>%
  left_join(
          data.frame(
                  language =top_langs,
                  lang_long = all_langs
          )
  ) |> select(-language) |> rename(language=lang_long) |> 
  group_by(gutenberg_id) %>%
  summarise(languages = list(sort(language))) %>%
  mutate(combo = map_chr(languages, ~paste(.x, collapse = "&"))) %>%
  count(combo, languages) %>%
  filter(n > 1)  

# Create presence/absence matrix

matrix_data <- df_upset %>%
  mutate(
    matrix = map(languages, function(langs) {
      setNames(as.list(as.integer(all_langs %in% langs)), all_langs)
    })
  ) %>%
  unnest_wider(matrix) %>%
  arrange(desc(n)) %>%
  mutate(combo_id = row_number())

# Reshape for plotting dots
dots_data <- matrix_data %>%
  select(combo_id, all_of(all_langs)) %>%
  pivot_longer(cols = all_of(all_langs), names_to = "language", values_to = "present") %>%
  filter(present == 1) %>%
  mutate(language = factor(language, levels = rev(all_langs)))

# Define colors
lang_colors <- setNames(
  brewer.pal(length(all_langs), "Paired"),
  all_langs
)

# Mix colors function
mix_colors <- function(langs) {
  if(length(langs) == 1) return(lang_colors[langs])
  cols <- col2rgb(lang_colors[langs])
  mixed <- rowMeans(cols)
  rgb(mixed[1], mixed[2], mixed[3], maxColorValue = 255)
}

# Add line colors
lines_data <- dots_data %>%
  group_by(combo_id) %>%
  summarise(
    y_min = min(as.numeric(language)),
    y_max = max(as.numeric(language)),
    langs = list(as.character(language)),
    line_color = mix_colors(unlist(langs))
  ) %>%
  filter(y_min != y_max)


```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}
# Matrix plot
p_matrix <- ggplot() +
  geom_segment(data = lines_data, 
               aes(x = combo_id, xend = combo_id, y = y_min, yend = y_max, color = line_color),
               linewidth = 3) +
  geom_point(data = dots_data, 
             aes(x = combo_id, y = as.numeric(language), color = language),
             size = 4) +
  scale_color_manual(values = c(lang_colors, setNames(lines_data$line_color, lines_data$line_color))) +
  scale_y_continuous(breaks = 1:length(all_langs), labels = rev(all_langs)) +
  scale_x_continuous(limits = c(0.5, max(matrix_data$combo_id) + 0.5), expand = c(0, 0)) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title = element_blank(),
        panel.grid.major.x = element_blank())

# Bar plot with log scale and limits
p_bars <- matrix_data %>%
  mutate(bar_color = map_chr(languages, ~mix_colors(.x))) %>%
  ggplot(aes(x = combo_id, y = n, fill = bar_color)) +
  geom_col(width = 1) +
  scale_fill_identity() +
  scale_y_log10(
    limits = c(1, 100000),
    labels = label_log(digits = 1),
    n.breaks=6
  ) +
  scale_x_continuous(limits = c(0.5, max(matrix_data$combo_id) + 0.5), expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())+
  ylab("Number of books")+
  ggtitle("Number of books per language and combinations")

plot_grid(p_bars, p_matrix, ncol = 1, align = "v", rel_heights = c(3, 2))

  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}
# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png"
)
```
